---
title: "Enrollment Report"
format: html
editor: visual
---

## Executive Summary

The 2025 Recruitment Cycle has been one of extensive innovation in regards to historic processes in place for GCE. Challenges have included less-than-stellar top of funnel numbers for our graduate programs, with new opportunities that have been capitalized on through a boon in our top of funnel numbers for our Accelerated Bachelor's Degree Completion pr

## Graduate Applications

Graduate Applications continue to lag behind

```{r}
#| label: Grad Applications
library(ggplot2)
library(dplyr)
library(lubridate)

# Load the CSV file
data <- `GradApps(2024.25).4.4.25`

# Ensure Application.Start.Date is in Date-Time format (MM/DD/YYYY)
data$Application.Start.Date <- mdy(data$Application.Start.Date)

# Extract month from Application.Start.Date
data$ApplicationMonth <- month(data$Application.Start.Date, label = TRUE)  

# Filter for Active.Term values (2024, 2025) and clean up the Year column
filtered_data <- data %>%
  filter(grepl("2024|2025", Active.Term...Calculated)) %>%
  mutate(Year = as.character(ifelse(grepl("2024", Active.Term...Calculated), "2024", "2025"))) %>%
  group_by(Active.Major...Calculated, Year) %>%
  summarise(count = n(), .groups = "drop")

# Ensure Active.Major...Calculated is a character column
filtered_data$Active.Major...Calculated <- as.character(filtered_data$Active.Major...Calculated)

# Create the bar chart
ggplot(filtered_data, aes(x = Active.Major...Calculated, y = count, fill = Year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Active Majors by Active Term (2024 & 2025)",
       x = "Active Major",
       y = "Count of Active Terms",
       fill = "Year") +
  scale_fill_manual(values = c("2024" = "blue", "2025" = "orange")) +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))  # Rotate x labels for readability



  


```

You can add options to executable code like this

```{r}
library(ggplot2)
library(dplyr)
library(lubridate)

# Load the CSV file
data <- Grad.Apps.2024.2025...muhlenberg.files.17387781

# Ensure Application.Start.Date is in Date-Time format (MM/DD/YYYY)
data$Application.Start.Date <- mdy(data$Application.Start.Date)

# Extract month from Application.Start.Date
data$ApplicationMonth <- month(data$Application.Start.Date, label = TRUE)  

# Filter for Active.Term values (2024, 2025) and clean up the Year column
filtered_data <- data %>%
  filter(grepl("2024|2025", Active.Term...Calculated)) %>%
  mutate(Year = as.character(ifelse(grepl("2024", Active.Term...Calculated), "2024", "2025"))) %>%
  group_by(Active.Major...Calculated, Year) %>%
  summarise(count = n(), .groups = "drop")

# Ensure Active.Major...Calculated is a character column
filtered_data$Active.Major...Calculated <- as.character(filtered_data$Active.Major...Calculated)

# Create the bar chart
ggplot(filtered_data, aes(x = Active.Major...Calculated, y = count, fill = Year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Active Majors by Active Term (2024 & 2025)",
       x = "Active Major",
       y = "Count of Active Terms",
       fill = "Year") +
  scale_fill_manual(values = c("2024" = "blue", "2025" = "orange")) +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))  # Rotate x labels for readability



  

```

The `echo: false` option disables the printing of code (only output is displayed).
