---
title: "Quarterly Enrollment Report"
format: 
  html:
    toc: true
    number-sections: true
  pdf:
    fig-width: 8
    fig-height: 5
    toc: true
    number-sections: true
toc: true
theme: cosmo
echo: false
---

# Prospects and Suspects (Top of Funnel)

## Executive Summary

No leads currently for 2026

```{r}
#| label: Leads
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
library(lubridate)
library(ggthemes)   # for theme_tufte()

# ---- Config ----
path <- "C:\\Users\\Todd\\Dropbox\\Monthly Export\\2025prospects_suspects.csv"

# Okabe-Ito palette
okabe_ito <- c("2023" = "#E69F00",  # orange
               "2024" = "#56B4E9",  # sky blue
               "2025" = "#009E73",  # bluish green
               "2026" = "#F0E442")  # yellow

# ---- Read & normalize columns ----
df <- readr::read_csv(path, show_col_types = FALSE) |>
  rename_with(make.names)  # turns spaces/dashes into dots

# Expect these normalized names based on your export:
#   Date.of.Inquiry
#   Active.Major...Calculated
stopifnot("Date.of.Inquiry" %in% names(df),
          "Active.Major...Calculated" %in% names(df))

# ---- Parse dates & derive time fields ----
df <- df |>
  mutate(
    Date = suppressWarnings(mdy(Date.of.Inquiry)),  # most of your files are m/d/Y
    # if you ever get NA's from odd rows, uncomment the fallback:
    # Date = if_else(is.na(Date), ymd(Date.of.Inquiry), Date),
    Year = year(Date),
    Month = month(Date, label = TRUE, abbr = TRUE, locale = "C"),
    Month_Num = month(Date)
  )

# ---- Filter months & years ----
df_q1 <- df |>
  filter(!is.na(Date)) |>
  filter(Month_Num %in% 1:8) |>
  filter(Year %in% c(2023, 2024, 2025, 2026))

# ---- Program type bucketing ----
df_q1 <- df_q1 |>
  mutate(Program_Type = case_when(
    str_detect(Active.Major...Calculated, regex("Accelerated Bachelor", ignore_case = TRUE)) ~ "Accelerated",
    str_detect(Active.Major...Calculated, regex("Master|Graduate Certificate", ignore_case = TRUE)) ~ "Graduate",
    TRUE ~ "All"
  ))

# ---- Summarize helper ----
summarize_leads <- function(d, program_type) {
  d |>
    filter(Program_Type == program_type | program_type == "All") |>
    group_by(Year, Month) |>
    summarise(Leads = n(), .groups = "drop") |>
    arrange(Year, Month) |>
    mutate(
      Cumulative_Leads = cumsum(Leads),
      Program_Type = program_type
    )
}

# ---- Build summaries ----
combined_summary <- bind_rows(
  summarize_leads(df_q1, "All"),
  summarize_leads(df_q1, "Accelerated"),
  summarize_leads(df_q1, "Graduate")
) |>
  mutate(
    Program_Type = factor(Program_Type, levels = c("All", "Accelerated", "Graduate")),
    Year = factor(Year)
  )

# ---- Plot ----
ggplot(combined_summary, aes(x = Month, y = Leads, group = Year, color = Year)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_wrap(~ Program_Type, ncol = 1) +
  labs(
    title = "Leads by Program Type (Jan–Aug)",
    y = "Number of Leads", x = "Month", color = "Year"
  ) +
  scale_color_manual(values = okabe_ito, drop = FALSE) +
  theme_tufte(base_size = 12) +
  theme(
    legend.position = "top",
    strip.text = element_text(face = "bold")
  )


```

# Applications

## School of Graduate Studies, Executive Summary

Graduate applications maintain some momentum despite evaporating top of funnel

### Cumulative Applications by Program

```{r}
#| echo: false
#| warning: false
#| message: false
library(ggplot2)
library(lubridate)
library(dplyr)
library(stringr)

# Define abbreviation mapping
all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL"
)

okabe_ito_10 <- c(
  "#E69F00",  # orange
  "#56B4E9",  # sky blue
  "#009E73",  # bluish green
  "#F0E442",  # yellow
  "#0072B2",  # blue
  "#D55E00",  # vermilion
  "#CC79A7",  # reddish purple
  "#000000",  # black
  "#999999",  # medium gray
  "#999933"   # olive/khaki
)


# Load the CSV file
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Gradapps20242025.csv")

# Clean the data
data <- data %>%
  mutate(
    Application.Start.Date = mdy(Application.Start.Date),
    Year = year(Application.Start.Date),
    Month = floor_date(Application.Start.Date, "month"),
    Major_Short = ifelse(
      as.character(Active.Major...Calculated) %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  ) %>%
  filter(!is.na(Application.Start.Date))

# Filter for 2025 graduate programs only
data_filtered <- data %>%
  filter(
    grepl("Graduate|Master", Active.Major...Calculated, ignore.case = TRUE) &
    grepl("2025", Active.Term...Calculated)
  )

# Aggregate and calculate cumulative counts
aggregated_data <- data_filtered %>%
  group_by(Month, Major_Short) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Major_Short) %>%
  mutate(cumulative_count = cumsum(count))

# Create the plot
ggplot(aggregated_data, aes(x = Month, y = cumulative_count, color = Major_Short, group = Major_Short)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = okabe_ito_10) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
  labs(title = "Cumulative Application Starts for Graduate Majors (2025)",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Major") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Applications by Program (YoY Comparison, with months )

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 8
#| out-width: 100%
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)

# Define abbreviation mapping
all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCHML",
  "Graduate Certificate in Digital Learning" = "GCDL"
)

# Load the CSV file
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Gradapps20242025.csv")

# Convert date field
data$Application.Start.Date <- mdy(data$Application.Start.Date)

# Extract filtered and labeled data
filtered_data <- data %>%
  filter(month(Application.Start.Date) %in% 1:9) %>%
  filter(grepl("2024|2025", Active.Term...Calculated)) %>%
  filter(!grepl("Accelerated", Active.Major...Calculated, ignore.case = TRUE)) %>%
  mutate(
    Year = ifelse(grepl("2024", Active.Term...Calculated), "2024", "2025"),
    Major_Short = ifelse(
      as.character(Active.Major...Calculated) %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)  # fallback in case of mismatch
    )
  ) %>%
  group_by(Major_Short, Year) %>%
  summarise(count = n(), .groups = "drop")

# Plot using Major_Short for clarity
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Application Starts (Jan-Aug, 2024 & 2025)",
       x = "Major",
       y = "Count of Active Terms",
       fill = "Year") +
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9")) +  # Colorblind-friendly
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))


```

### Application Submitted by Program, Quarterly Analysis

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 8
#| out-width: 100%
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)
library(forcats)
library(ggthemes)

# Abbreviation mapping
all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  # ADCP
  "Accelerated Bachelor's in Business Administration"      = "Business Admin",
  "Accelerated Bachelor's in Business Analytics"           = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity"                = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management"        = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership"   = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management"           = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management"      = "Supply Chain"
)

# Load
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Gradapps20242025.csv")

# Parse submit/decision date (robust to time stamps)
data$Decision.created <- parse_date_time(
  data$Decision.created,
  orders = c("mdy HMS","mdy HM","mdy","ymd HMS","ymd HM","ymd"),
  tz = "UTC",
  quiet = TRUE
)

# Filter + label + abbreviate
filtered_data <- data %>%
  filter(!is.na(Decision.created), month(Decision.created) %in% 1:8) %>%
  filter(grepl("2024|2025", Active.Term...Calculated)) %>%
  # Keep graduate portfolio only (exclude ADCP)
  filter(!grepl("Accelerated", Active.Major...Calculated, ignore.case = TRUE)) %>%
  mutate(
    Year = ifelse(grepl("2024", Active.Term...Calculated), "2024", "2025"),
    Major_Short = ifelse(
      is.na(Active.Major...Calculated) | Active.Major...Calculated == "",
      "No Major",
      ifelse(
        Active.Major...Calculated %in% names(all_abbreviations),
        all_abbreviations[Active.Major...Calculated],
        Active.Major...Calculated
      )
    )
  ) %>%
  group_by(Major_Short, Year) %>%
  summarise(count = n(), .groups = "drop")

# Order majors by total across both years (desc)
order_levels <- filtered_data %>%
  group_by(Major_Short) %>%
  summarise(total = sum(count), .groups = "drop") %>%
  arrange(desc(total)) %>%
  pull(Major_Short)

filtered_data <- filtered_data %>%
  mutate(Major_Short = factor(Major_Short, levels = order_levels))

# Plot
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  labs(
    title = "Graduate Applications Submitted by Active Term (Jan–Aug, 2024 & 2025)",
    x = "Major",
    y = "Count of Applications Submitted",
    fill = "Year"
  ) +
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9")) +  # Okabe–Ito
  theme_tufte(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = "top",
    plot.title.position = "plot"
  )

```

### Application Start Distribution, year over year

```{r}
#| echo: false
#| warning: false
#| message: false

library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)
library(ggthemes)  # for theme_tufte()

# ---- Load CSV ----
# (Use your existing file; this will gracefully include 2026 rows if present)
data <- read_csv("C:/Users/Todd/Dropbox/Monthly Export/Gradapps20242025.csv",
                 show_col_types = FALSE)

# ---- Parse dates & derive year from Start Term ----
data <- data %>%
  mutate(
    # robust date parsing
    `Application.Start.Date` = suppressWarnings(mdy(`Application Start Date`)),
    # pull the first 4-digit year token from Active.Term...Calculated
    Year = str_extract(`Active Term - Calculated`, "\\b20\\d{2}\\b")
  ) %>%
  # keep only target years
  filter(Year %in% c("2024", "2025", "2026"))

# ---- Optional daily counts (kept in case you need it later) ----
daily_counts <- data %>%
  filter(!is.na(`Application.Start.Date`)) %>%
  group_by(`Application.Start.Date`, Year) %>%
  summarise(Count = n(), .groups = "drop")

# ---- Okabe–Ito palette ----
okabe_ito <- c(
  "2024" = "#E69F00",  # orange
  "2025" = "#56B4E9",  # sky blue
  "2026" = "#009E73"   # bluish green (your requested green)
)

# ---- Plot (histogram over time, monthly-ish bins) ----
p <- ggplot(
  data %>% filter(!is.na(`Application.Start.Date`)),
  aes(x = `Application.Start.Date`, fill = Year)
) +
  geom_histogram(binwidth = 30, alpha = 0.7, position = "identity", boundary = as.Date("2024-01-01")) +
  scale_x_date(date_breaks = "3 months", date_labels = "%m-%y") +
  scale_fill_manual(values = okabe_ito, breaks = c("2024", "2025", "2026")) +
  labs(
    title = "Graduate Application Start Distribution (2024–2026)",
    x = "Application Start Date",
    y = "Number of Applications Started",
    fill = "Start Term Year"
  ) +
  theme_tufte(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    panel.grid.major = element_line(linewidth = 0.2, linetype = "dotted")
  )

print(p)


```

### Cumulative Starts by year and program, small multiples

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
# Load required libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)

all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCHML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  
  # ADCP
  "Accelerated Bachelor's in Business Administration" = "Business Admin",
  "Accelerated Bachelor's in Business Analytics" = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity" = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management" = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership" = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management" = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management" = "Supply Chain"
)

data <- read.csv("data/Gradapps242526.csv")
colnames(data)
# Clean the data
data <- data %>%
  mutate(Application.Start.Date = mdy(Application.Start.Date),  # Convert MM/DD/YYYY to Date format
         Year = year(Application.Start.Date),  # Extract Year
         Month = floor_date(Application.Start.Date, "month"),  # Keep Month
         Active.Major...Calculated = as.factor(Active.Major...Calculated)) %>%
  filter(!is.na(Application.Completed.Date))  # Remove rows with NA values



# Check if Active.Term...Calculated exists before using it
if ("Active.Term...Calculated" %in% colnames(data)) {
  data <- data %>%
    mutate(Active.Term.Year = str_extract(as.character(Active.Term...Calculated), "\\d{4}"))  # Extract Year
} else {
  stop("Error: 'Active.Term...Calculated' column not found in the dataset.")
}

# 🔹 Filter for Graduate/Master's programs and select both 2024 & 2025 terms
data_filtered <- data %>%
  filter(grepl("Graduate|Master", Active.Major...Calculated, ignore.case = TRUE) &
           Active.Term.Year %in% c("2024", "2025", "2026"))  # Include both years

# Aggregate data by Month, Major, and Year, then compute cumulative sum
aggregated_data <- data_filtered %>%
  group_by(Month, Active.Major...Calculated, Active.Term.Year) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Active.Major...Calculated, Active.Term.Year) %>%
  mutate(cumulative_count = cumsum(count))  # Compute cumulative sum

# program name abbreviations 
aggregated_data <- aggregated_data %>%
  mutate(
    Major_Short = ifelse(
      as.character(Active.Major...Calculated) %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  )




data <- data %>%
  mutate(
    Major_Short = ifelse(
      Active.Major...Calculated %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  )


# Create the small multiples plot
plot <- ggplot(aggregated_data, aes(x = Month, y = cumulative_count, 
                            color = Active.Term.Year, 
                            group = Active.Term.Year)) +
  geom_point(size = 2) + 
  geom_line(size = 1) +
  scale_color_brewer(palette = "Set2") +  
  scale_x_date(date_breaks = "6 months", date_labels = "%b-%Y") +  
  labs(title = "Cumulative Application Starts by Major & Year",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Application Year") +  # Updated legend title
  facet_wrap(~ Major_Short, scales = "free_y") +  # Small multiples
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(plot)

```

## Accelerated Degree Completion Program (ADCP), Executive Summary

### Cumulative Applications by Program

```{r}
#| echo: false
#| warning: false
#| message: false
library(ggplot2)
library(lubridate)
library(dplyr)

# Read the CSV data
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv")  # Replace with your actual CSV file path

# Clean the data
data <- data %>%
  mutate(Application.Start.Date = mdy(Application.Start.Date),  # Convert MM/DD/YYYY to Date format
         Year = year(Application.Start.Date),
         Month = floor_date(Application.Start.Date, "month"),  # Group by month
         Active.Major...Calculated = as.factor(Active.Major...Calculated)) %>%
  filter(!is.na(Application.Start.Date))  # Remove rows with NA values in Application.Start.Date

# 🔹 Ensure you filter by both Major and Entry Term BEFORE aggregation
data_filtered <- data %>%
  filter(
    grepl("2025", Active.Term...Calculated),
    !grepl("^(Graduate|Master)", Active.Major...Calculated)
  )  # Change "2025" for other terms as needed

# Aggregate data by Month and Major, then compute cumulative sum
aggregated_data <- data_filtered %>%
  group_by(Month, Active.Major...Calculated) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Active.Major...Calculated) %>%
  mutate(cumulative_count = cumsum(count))  # Compute cumulative sum **AFTER filtering**

library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove(Active.Major...Calculated, "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in "))

# Now use Major_Short for labeling
ggplot(aggregated_data, aes(x = Month, y = cumulative_count, color = Major_Short, group = Major_Short)) +
  geom_line(size = 1) +  
  geom_point(size = 2) +
  scale_color_manual(values = okabe_ito_10) +   
  scale_x_date(date_breaks = "3 months", date_labels = "%m-%Y") +  
  labs(title = "Cumulative Application Starts for ADCP Majors (2025)",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Concentration") +  # updated legend title
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Applications by Program

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)
library(forcats)
library(ggthemes)

# Abbreviation mapping
all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  # ADCP
  "Accelerated Bachelor's in Business Administration"        = "Business Admin",
  "Accelerated Bachelor's in Business Analytics"             = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity"                  = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management"          = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership"     = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management"             = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management"        = "Supply Chain"
)

# Load
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv", stringsAsFactors = FALSE)

# Parse Application.Start.Date (robust to timestamps)
data$Application.Start.Date <- parse_date_time(
  data$Application.Start.Date,
  orders = c("mdy HMS","mdy HM","mdy","ymd HMS","ymd HM","ymd"),
  tz = "UTC", quiet = TRUE
)

# Filter + label + abbreviate
filtered_data <- data %>%
  filter(!is.na(Application.Start.Date), month(Application.Start.Date) %in% 1:8) %>%
  filter(str_detect(Active.Term...Calculated, "2024|2025")) %>%
  # Keep accelerated portfolio only (exclude Graduate/Master)
  filter(!str_detect(Active.Major...Calculated, regex("Graduate|Master", ignore_case = TRUE))) %>%
  mutate(
    Year = if_else(str_detect(Active.Term...Calculated, "2024"), "2024", "2025"),
    Major_Short = case_when(
      is.na(Active.Major...Calculated) | Active.Major...Calculated == "" ~ "No Major",
      Active.Major...Calculated %in% names(all_abbreviations) ~ all_abbreviations[Active.Major...Calculated],
      TRUE ~ Active.Major...Calculated
    )
  ) %>%
  group_by(Major_Short, Year) %>%
  summarise(count = n(), .groups = "drop")

# Order majors by total across both years (desc)
order_levels <- filtered_data %>%
  group_by(Major_Short) %>%
  summarise(total = sum(count), .groups = "drop") %>%
  arrange(desc(total)) %>%
  pull(Major_Short)

filtered_data <- filtered_data %>%
  mutate(Major_Short = factor(Major_Short, levels = order_levels))

# Plot
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  labs(
    title = "Accelerated Applications Started by Active Term (Jan–Aug, 2024 & 2025)",
    x = "Major",
    y = "Count of Application Starts",
    fill = "Year"
  ) +
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9")) +  # Okabe–Ito
  theme_tufte(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = "top",
    plot.title.position = "plot"
  )

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"

library(tidyverse)
library(lubridate)
library(ggthemes)   # theme_tufte()

# ---- Config ----
path <- "C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv"
okabe_ito <- c("2024" = "#E69F00", "2025" = "#56B4E9")

# Abbreviation mapping
all_abbreviations <- c(
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  "Accelerated Bachelor's in Business Administration" = "Business Admin",
  "Accelerated Bachelor's in Business Analytics" = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity" = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management" = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership" = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management" = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management" = "Supply Chain"
)

# ---- Read & normalize ----
dat <- readr::read_csv(path, show_col_types = FALSE) |>
  rename_with(make.names)

# Required columns after make.names()
req <- c("Active.Term...Calculated", "Active.Major...Calculated", "Application.Completed.Date")
missing <- setdiff(req, names(dat))
if (length(missing) > 0) {
  stop("Missing required columns: ", paste(missing, collapse = ", "))
}

# Robust parser that handles m/d/Y, Y-m-d, and ISO with time ('T...Z')
parse_any <- function(x) {
  # try m/d/Y
  d1 <- suppressWarnings(mdy(x))
  # try plain Y-m-d
  d2 <- suppressWarnings(ymd(x))
  # try ISO datetime
  d3 <- suppressWarnings(ymd_hms(x, quiet = TRUE))
  out <- coalesce(d1, d2, as_date(d3))
  as_date(out)
}

# ---- Parse & filter Jan–Aug by Completed Date ----
dat <- dat |>
  mutate(Date_for_Month = parse_any(Application.Completed.Date)) |>
  filter(!is.na(Date_for_Month)) |>
  filter(month(Date_for_Month) %in% 1:8)

# ---- Keep 2024/2025 and Accelerated (exclude Graduate/Master) ----
filtered_data <- dat |>
  filter(str_detect(Active.Term...Calculated, "2024|2025")) |>
  filter(!str_detect(Active.Major...Calculated, regex("Graduate|Master", ignore_case = TRUE))) |>
  mutate(
    Year = if_else(str_detect(Active.Term...Calculated, "2024"), "2024", "2025"),
    Major_Short = if_else(
      as.character(Active.Major...Calculated) %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  ) |>
  group_by(Major_Short, Year) |>
  summarise(count = n(), .groups = "drop")

if (nrow(filtered_data) == 0) {
  stop("No rows after filtering. Confirm there are Jan–Aug completed dates and 2024/2025 in Active.Term...Calculated.")
}

# ---- Order majors by total across years (desc) ----
order_levels <- filtered_data |>
  group_by(Major_Short) |>
  summarise(total = sum(count), .groups = "drop") |>
  arrange(desc(total)) |>
  pull(Major_Short)

filtered_data <- filtered_data |>
  mutate(
    Major_Short = factor(Major_Short, levels = order_levels),
    Year = factor(Year, levels = c("2024","2025"))
  )

# ---- Plot ----
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  labs(
    title = "Accelerated Applications Completed (Jan–Aug) by Major\n(Source: Application Completed Date)",
    x = "Major",
    y = "Count",
    fill = "Year"
  ) +
  scale_fill_manual(values = okabe_ito, drop = FALSE) +
  theme_tufte(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = "top"
  )



```

### Application Start Distribution, year over year

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 8
#| out-width: 100%
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)

# Load CSV
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv")

# Convert Application Start Date to date format
data <- data %>%
  mutate(Application.Start.Date = as.Date(Application.Start.Date, format = "%m/%d/%Y"))

# Extract year from Start Term and filter for 2024 or 2025
data_filtered <- data %>%
  filter(grepl("2024|2025", Active.Term...Calculated)) %>%
  mutate(Year = ifelse(grepl("2024", Active.Term...Calculated), "2024", "2025"))

# Aggregate data: count application starts per day for each year
daily_counts <- data_filtered %>%
  group_by(Application.Start.Date, Year) %>%
  summarise(Count = n(), .groups = "drop")
library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove(Active.Major...Calculated, "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in "))

plot <- ggplot(data_filtered, aes(x = Application.Start.Date, fill = Year)) +
  geom_histogram(binwidth = 30, alpha = 0.6, position = "identity") +  # Semi-transparent overlap
  scale_color_brewer(palette = "Set2") +
  scale_x_date(date_breaks = "3 months", date_labels = "%m-%y") +  # Show full month and year
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9")) +
  labs(title = "Application Start Distribution (2024 vs. 2025)",
       x = "Application Start Date",
       y = "Number of Applications Started",
       fill = "Start Term Year") +
  theme_minimal(base_size=10) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot)
```

### Cumulative Starts by year and program, small multiples

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
# Load required libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)

data <- read.csv("data/AcceleratedApps.csv")

# Clean the data
data <- data %>%
  mutate(Application.Start.Date = mdy(Application.Start.Date),  # Convert MM/DD/YYYY to Date format
         Year = year(Application.Start.Date),  # Extract Year
         Month = floor_date(Application.Start.Date, "month"),  # Keep Month
         Active.Major...Calculated = as.factor(Active.Major...Calculated)) %>%
  filter(!is.na(Application.Start.Date))  # Remove rows with NA values

# Check if Active.Term...Calculated exists before using it
if ("Active.Term...Calculated" %in% colnames(data)) {
  data <- data %>%
    mutate(Active.Term.Year = str_extract(as.character(Active.Term...Calculated), "\\d{4}"))  # Extract Year
} else {
  stop("Error: 'Active.Term...Calculated' column not found in the dataset.")
}

# 🔹 Filter for Graduate/Master's programs and select both 2024 & 2025 terms
data_filtered <- data %>%
  filter(grepl("Accelerated", Active.Major...Calculated, ignore.case = TRUE) &
           Active.Term.Year %in% c("2024", "2025"))  # Include both years

# Aggregate data by Month, Major, and Year, then compute cumulative sum
aggregated_data <- data_filtered %>%
  group_by(Month, Active.Major...Calculated, Active.Term.Year) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Active.Major...Calculated, Active.Term.Year) %>%
  mutate(cumulative_count = cumsum(count))  # Compute cumulative sum

library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove(Active.Major...Calculated, "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in ") %>% str_remove("Master"))

# Create the small multiples plot
plot <- ggplot(aggregated_data, aes(x = Month, y = cumulative_count, 
                            color = Active.Term.Year, 
                            group = Active.Term.Year)) +
  geom_point(size = 2) + 
  geom_line(size = 1) +
  scale_color_brewer(palette = "Set2") +  
  scale_x_date(date_breaks = "6 months", date_labels = "%b-%Y") +  
  labs(title = "Cumulative Application Starts by Major & Year",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Application Year") +  # Updated legend title
  facet_wrap(~ Major_Short, scales = "free_y") +  # Small multiples
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(plot)

```

# Analysis
