------
title: "Quarterly Enrollment Report"
format: 
  html:
    toc: true
    number-sections: true
  pdf:
    fig-width: 8
    fig-height: 5
    toc: true
    number-sections: true
toc: true
theme: cosmo
echo: false
---

# Prospects and Suspects (Top of Funnel)

## Executive Summary

No leads currently for 2026

```{r}
suppressPackageStartupMessages({
  library(tidyverse); library(lubridate); library(ggthemes)
})

path <- "C:\\Users\\Todd\\Dropbox\\Monthly Export\\2025prospects_suspects.csv"
as_of <- as.Date("2025-11-03")  # lock for reproducibility
okabe_ito <- c("2024"="#56B4E9","2025"="#009E73","2026"="#F0E442")

df <- readr::read_csv(path, show_col_types = FALSE) |>
  rename_with(make.names) |>
  mutate(
    Date = coalesce(suppressWarnings(mdy(Date.of.Inquiry)),
                    suppressWarnings(ymd(Date.of.Inquiry))),
    Entry_Year = str_extract(Active.Term...Calculated, "(20\\d{2})") |> as.integer()
  )

# Program type buckets (Accelerated & Graduate only)
df <- df |>
  mutate(
    Program_Type = case_when(
      str_detect(Active.Student.Type...Calculated, regex("accelerated", TRUE)) |
      str_detect(Active.Major...Calculated, regex("accelerated", TRUE)) ~ "Accelerated",
      str_detect(Active.Student.Type...Calculated, regex("graduate", TRUE)) |
      str_detect(Active.Major...Calculated, regex("master|graduate certificate", TRUE)) ~ "Graduate",
      str_detect(Active.Student.Type...Calculated, regex("continuing", TRUE)) ~ "Continuing",
      str_detect(Active.Student.Type...Calculated, regex("non.?credit", TRUE)) ~ "Non-Credit",
      TRUE ~ "Other"
    )
  ) |>
  filter(Program_Type %in% c("Accelerated","Graduate"),
         !is.na(Date), !is.na(Entry_Year))

# Keep inquiries within Aug 1 -> same mm-dd as `as_of` in the recruiting year, dedupe by Email+Entry_Year
df_win <- df |>
  mutate(
    Recruit_Year = Entry_Year - 1L,
    start_win = make_date(Recruit_Year, 8, 1),
    end_win   = make_date(Recruit_Year, month(as_of), mday(as_of))
  ) |>
  filter(Date >= start_win, Date <= end_win) |>
  arrange(Date) |>
  distinct(Email, Entry_Year, .keep_all = TRUE)

# Totals
totals <- df_win |>
  count(Entry_Year, name = "Inquiries_to_Date") |>
  arrange(Entry_Year)

print(totals)

# Optional: grouped by program type
by_type <- df_win |>
  count(Entry_Year, Program_Type, name = "Inquiries_to_Date") |>
  arrange(Entry_Year, Program_Type)

print(by_type)

# Plot
ggplot(totals, aes(x=factor(Entry_Year), y=Inquiries_to_Date, fill=factor(Entry_Year))) +
  geom_col() +
  geom_text(aes(label=Inquiries_to_Date), vjust=-0.25, size=4) +
  scale_fill_manual(values=okabe_ito) +
  labs(
    title = paste0("Cumulative inquiries (Aug 1–", format(as_of, "%b %d"), ") by Entry Term"),
    x = "Entry Term Year", y = "Unique inquiries to date"
  ) +
  theme_tufte(base_size = 12) +
  theme(legend.position = "none")


```

# Applications

## School of Graduate Studies, Executive Summary

Graduate applications maintain some momentum despite evaporating top of funnel

### Cumulative Applications by Program

```{r}
#| echo: false
#| warning: false
#| message: false
library(ggplot2)
library(lubridate)
library(dplyr)
library(stringr)

# Define abbreviation mapping
all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  "Master's in Business Administration" = "MBA",
  
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL"
)

okabe_ito_10 <- c(
  "#E69F00",  # orange
  "#56B4E9",  # sky blue
  "#009E73",  # bluish green
  "#F0E442",  # yellow
  "#0072B2",  # blue
  "#D55E00",  # vermilion
  "#CC79A7",  # reddish purple
  "#000000",  # black
  "#999999",  # medium gray
  "#999933"   # olive/khaki
)


# Load the CSV file
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Gradapps20242025.csv")

# Clean the data
data <- data %>%
  mutate(
    Application.Start.Date = mdy(Application.Start.Date),
    Year = year(Application.Start.Date),
    Month = floor_date(Application.Start.Date, "month"),
    Major_Short = ifelse(
      as.character(Application.Major) %in% names(all_abbreviations),
      all_abbreviations[as.character(Application.Major)],
      as.character(Application.Major)
    )
  ) %>%
  filter(!is.na(Application.Start.Date))

# Filter for 2025 graduate programs only
data_filtered <- data %>%
  filter(
    grepl("Graduate|Master", Application.Major, ignore.case = TRUE) &
    grepl("2026", Active.Term...Calculated)
  )

# Aggregate and calculate cumulative counts
aggregated_data <- data_filtered %>%
  group_by(Month, Major_Short) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Major_Short) %>%
  mutate(cumulative_count = cumsum(count))

# Create the plot
ggplot(aggregated_data, aes(x = Month, y = cumulative_count, color = Major_Short, group = Major_Short)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = okabe_ito_10) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b-%Y") +
  labs(title = "Cumulative Application Starts for Graduate Majors (2026)",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Major") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Applications by Program (YoY Comparison, with months )

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 8
#| out-width: 100%
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)

# Load CSV
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Gradapps20242025.csv")

# Convert Application Start Date to date format
data <- data %>%
  mutate(Application.Start.Date = as.Date(Application.Start.Date, format = "%m/%d/%Y"))

# Extract year from Start Term and filter for 2024 or 2025
data_filtered <- data %>%
  filter(grepl("2024|2025|2026", Active.Term...Calculated)) %>%
   mutate(
    Year = case_when(
      grepl("2024", Active.Term...Calculated) ~ "2024",
      grepl("2025", Active.Term...Calculated) ~ "2025",
      grepl("2026", Active.Term...Calculated) ~ "2026",
      TRUE ~ "Other"))

# Aggregate data: count application starts per day for each year
daily_counts <- data_filtered %>%
  group_by(Application.Start.Date, Year) %>%
  summarise(Count = n(), .groups = "drop")
library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove("Active.Major...Calculated", "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in "))

plot <- ggplot(data_filtered, aes(x = Application.Start.Date, fill = Year)) +
  geom_histogram(binwidth = 30, alpha = 0.6, position = "identity") +  # Semi-transparent overlap
  scale_color_brewer(palette = "Set2") +
  scale_x_date(date_breaks = "3 months", date_labels = "%m-%y") +  # Show full month and year
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9", "2026" = "#009E73")) +
  labs(title = "Application Start Distribution",
       x = "Application Start Date",
       y = "Number of Applications Started",
       fill = "Start Term Year") +
  theme_minimal(base_size=10) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot)


```

### Application Submitted by Program, Quarterly Analysis

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(lubridate)
  library(stringr)
  library(tidyr)
  library(ggplot2)
  library(ggthemes)
})
colnames(data)
# --- Window config (Aug–Nov) ---
start_month   <- 8
window_months <- 4
allowed_months <- ((start_month - 1 + 0:(window_months - 1)) %% 12) + 1  # c(8,9,10,11)

# Parse SUBMISSION date (robust to timestamps)
data <- data %>%
  mutate(
    Submitted_DT = parse_date_time(
      data$Application.Start.Date, 
      orders = c("mdy HMS","mdy HM","mdy","ymd HMS","ymd HM","ymd"),
      tz = "UTC",
      quiet = TRUE
    )
  )

# ---- Filter to Aug–Nov submissions; Grad only; label years ----
filtered_data <- data %>%
  filter(!is.na(Submitted_DT)) %>%
  filter(month(Submitted_DT) %in% allowed_months) %>%
  # keep only 2024/2025/2026 active terms
  filter(str_detect(Active.Term...Calculated, "2024|2025|2026")) %>%
  # exclude Accelerated/ADCP by major name (safer than a missing column)
  filter(!str_detect(Application.Major, regex("^Accelerated|Bachelor|Biology", ignore_case = TRUE))) %>%
  mutate(
    Year = case_when(
      str_detect(Active.Term...Calculated, "2024") ~ "2024",
      str_detect(Active.Term...Calculated, "2025") ~ "2025",
      str_detect(Active.Term...Calculated, "2026") ~ "2026",
      TRUE ~ "Other"
    ),
    Major_Short = case_when(
      is.na(Application.Major) | Application.Major == "" ~ "No Major",
      Application.Major %in% names(all_abbreviations) ~ unname(all_abbreviations[Application.Major]),
      TRUE ~ `Application.Major`
    )
  ) %>%
  group_by(Major_Short, Year) %>%
  summarise(count = n(), .groups = "drop")

# --- ensure all grad abbreviations (incl. MBA) appear even if zero ---
grad_abbrevs <- unname(all_abbreviations[!grepl("^Accelerated", names(all_abbreviations))])

filtered_data <- filtered_data %>%
  tidyr::complete(
    Major_Short = grad_abbrevs,
    Year = c("2024","2025","2026"),
    fill = list(count = 0)
  )

# Order majors by total across years
order_levels <- filtered_data %>%
  group_by(Major_Short) %>%
  summarise(total = sum(count), .groups = "drop") %>%
  arrange(desc(total)) %>%
  pull(Major_Short)

filtered_data <- filtered_data %>%
  mutate(Major_Short = factor(Major_Short, levels = order_levels))

# Plot (Okabe–Ito + Tufte)
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  labs(
    title = "Graduate Applications Submitted (Aug–Nov window) by Active Term",
    x = "Major",
    y = "Count of Applications Submitted",
    fill = "Year"
  ) +
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9", "2026" = "#009E73")) +
  theme_tufte(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = "top",
    plot.title.position = "plot"
  )




```

### Cumulative Starts by year and program, small multiples

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
# Load required libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)

all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCHML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  
  # ADCP
  "Accelerated Bachelor's in Business Administration" = "Business Admin",
  "Accelerated Bachelor's in Business Analytics" = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity" = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management" = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership" = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management" = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management" = "Supply Chain"
)

data <- read.csv("data/Gradapps242526.csv")

# Clean the data
data <- data %>%
  mutate(Application.Start.Date = mdy(Application.Start.Date),  # Convert MM/DD/YYYY to Date format
         Year = year(Application.Start.Date),  # Extract Year
         Month = floor_date(Application.Start.Date, "month"),  # Keep Month
         Active.Major...Calculated = as.factor(Active.Major...Calculated)) %>%
  filter(!is.na(Application.Completed.Date))  # Remove rows with NA values



# Check if Active.Term...Calculated exists before using it
if ("Active.Term...Calculated" %in% colnames(data)) {
  data <- data %>%
    mutate(Active.Term.Year = str_extract(as.character(Active.Term...Calculated), "\\d{4}"))  # Extract Year
} else {
  stop("Error: 'Active.Term...Calculated' column not found in the dataset.")
}

# 🔹 Filter for Graduate/Master's programs and select both 2024 & 2025 terms
data_filtered <- data %>%
  filter(grepl("Graduate|Master", Active.Major...Calculated, ignore.case = TRUE) &
           Active.Term.Year %in% c("2024", "2025", "2026"))  # Include both years

# Aggregate data by Month, Major, and Year, then compute cumulative sum
aggregated_data <- data_filtered %>%
  group_by(Month, Active.Major...Calculated, Active.Term.Year) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Active.Major...Calculated, Active.Term.Year) %>%
  mutate(cumulative_count = cumsum(count))  # Compute cumulative sum

# program name abbreviations 
aggregated_data <- aggregated_data %>%
  mutate(
    Major_Short = ifelse(
      as.character(Active.Major...Calculated) %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  )




data <- data %>%
  mutate(
    Major_Short = ifelse(
      Active.Major...Calculated %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  )


# Create the small multiples plot
plot <- ggplot(aggregated_data, aes(x = Month, y = cumulative_count, 
                            color = Active.Term.Year, 
                            group = Active.Term.Year)) +
  geom_point(size = 2) + 
  geom_line(size = 1) +
  scale_color_brewer(palette = "Set2") +  
  scale_x_date(date_breaks = "6 months", date_labels = "%b-%Y") +  
  labs(title = "Cumulative Application Starts by Major & Year",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Application Year") +  # Updated legend title
  facet_wrap(~ Major_Short, scales = "free_y") +  # Small multiples
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(plot)

```

## Accelerated Degree Completion Program (ADCP), Executive Summary

### Cumulative Applications by Program

```{r}
#| echo: false
#| warning: false
#| message: false
library(ggplot2)
library(lubridate)
library(dplyr)

# Read the CSV data
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv")  # Replace with your actual CSV file path

# Clean the data
data <- data %>%
  mutate(Application.Start.Date = mdy(Application.Start.Date),  # Convert MM/DD/YYYY to Date format
         Year = year(Application.Start.Date),
         Month = floor_date(Application.Start.Date, "month"),  # Group by month
         Active.Major...Calculated = as.factor(Active.Major...Calculated)) %>%
  filter(!is.na(Application.Start.Date))  # Remove rows with NA values in Application.Start.Date

# 🔹 Ensure you filter by both Major and Entry Term BEFORE aggregation
data_filtered <- data %>%
  filter(
    grepl("2026", Active.Term...Calculated),
    !grepl("^(Graduate|Master)", Active.Major...Calculated)
  )  # Change "2025" for other terms as needed

# Aggregate data by Month and Major, then compute cumulative sum
aggregated_data <- data_filtered %>%
  group_by(Month, Active.Major...Calculated) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Active.Major...Calculated) %>%
  mutate(cumulative_count = cumsum(count))  # Compute cumulative sum **AFTER filtering**

library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove(Active.Major...Calculated, "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in "))

# Now use Major_Short for labeling
ggplot(aggregated_data, aes(x = Month, y = cumulative_count, color = Major_Short, group = Major_Short)) +
  geom_line(size = 1) +  
  geom_point(size = 2) +
  scale_color_manual(values = okabe_ito_10) +   
  scale_x_date(date_breaks = "3 months", date_labels = "%m-%Y") +  
  labs(title = "Cumulative Application Starts for ADCP Majors (2025)",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Concentration") +  # updated legend title
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Applications by Program

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
library(ggplot2)
library(dplyr)
library(lubridate)
library(stringr)
library(forcats)
library(ggthemes)

# Abbreviation mapping
all_abbreviations <- c(
  # Master's
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  # Graduate Certificates
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  # ADCP
  "Accelerated Bachelor's in Business Administration"        = "Business Admin",
  "Accelerated Bachelor's in Business Analytics"             = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity"                  = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management"          = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership"     = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management"             = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management"        = "Supply Chain"
)

# Load
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv", stringsAsFactors = FALSE)

# Parse Application.Start.Date (robust to timestamps)
data$Application.Start.Date <- parse_date_time(
  data$Application.Start.Date,
  orders = c("mdy HMS","mdy HM","mdy","ymd HMS","ymd HM","ymd"),
  tz = "UTC", quiet = TRUE
)

# Filter + label + abbreviate
filtered_data <- data %>%
  filter(!is.na(Application.Start.Date), month(Application.Start.Date) %in% 8:11) %>%
  filter(str_detect(Active.Term...Calculated, "2024|2025|2026")) %>%
  # Keep accelerated portfolio only (exclude Graduate/Master)
  filter(!str_detect(Active.Major...Calculated, regex("Graduate|Master", ignore_case = TRUE))) %>%
   mutate(
    Year = case_when(
      grepl("2024", Active.Term...Calculated) ~ "2024",
      grepl("2025", Active.Term...Calculated) ~ "2025",
      grepl("2026", Active.Term...Calculated) ~ "2026",
      TRUE ~ "Other"
    ),
    Major_Short = case_when(
      is.na(Active.Major...Calculated) | Active.Major...Calculated == "" ~ "No Major",
      Active.Major...Calculated %in% names(all_abbreviations) ~ all_abbreviations[Active.Major...Calculated],
      TRUE ~ Active.Major...Calculated
    )
  ) %>%
  group_by(Major_Short, Year) %>%
  summarise(count = n(), .groups = "drop")

# Order majors by total across both years (desc)
order_levels <- filtered_data %>%
  group_by(Major_Short) %>%
  summarise(total = sum(count), .groups = "drop") %>%
  arrange(desc(total)) %>%
  pull(Major_Short)

filtered_data <- filtered_data %>%
  mutate(Major_Short = factor(Major_Short, levels = order_levels))

# Plot
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  labs(
    title = "Accelerated Applications Started by Active Term (Aug–Nov, 2024 - 2026)",
    x = "Major",
    y = "Count of Application Starts",
    fill = "Year"
  ) +
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9", "2026" = "#009E73")) +  # Okabe–Ito
  theme_tufte(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = "top",
    plot.title.position = "plot"
  )

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"

library(tidyverse)
library(lubridate)
library(ggthemes)   # theme_tufte()

# ---- Config ----
path <- "C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv"
okabe_ito <- c("2024" = "#E69F00", "2025" = "#56B4E9", "2026" = "#009E73")

# Abbreviation mapping
all_abbreviations <- c(
  "Master's in Applied Analytics" = "MAA",
  "Master's in Organizational Leadership" = "MOL",
  "Master's in Healthcare Management and Leadership" = "MML",
  "Master's in Digital Learning" = "MDL",
  "Master of Arts in Teaching" = "MAT",
  "Graduate Certificate in Diversity, Equity, Inclusion and Belonging" = "DEIB",
  "Graduate Certificate in Organizational Leadership" = "GCOL",
  "Graduate Certificate in Applied Analytics" = "GCAA",
  "Graduate Certificate in Healthcare Management and Leadership" = "GCMML",
  "Graduate Certificate in Digital Learning" = "GCDL",
  "Accelerated Bachelor's in Business Administration" = "Business Admin",
  "Accelerated Bachelor's in Business Analytics" = "Bus. Analytics",
  "Accelerated Bachelor's in Cybersecurity" = "Cybersecurity",
  "Accelerated Bachelor's in Healthcare Management" = "Healthcare Mgmt",
  "Accelerated Bachelor's in Human Resources Leadership" = "HR Leadership",
  "Accelerated Bachelor's in Information Systems Management" = "Info Sys Mgmt",
  "Accelerated Bachelor's in Project Management" = "Project Mgmt",
  "Accelerated Bachelor's in Supply Chain Management" = "Supply Chain"
)

# ---- Read & normalize ----
dat <- readr::read_csv(path, show_col_types = FALSE) |>
  rename_with(make.names)

# Required columns after make.names()
req <- c("Active.Term...Calculated", "Active.Major...Calculated", "Application.Completed.Date")
missing <- setdiff(req, names(dat))
if (length(missing) > 0) {
  stop("Missing required columns: ", paste(missing, collapse = ", "))
}

# Robust parser that handles m/d/Y, Y-m-d, and ISO with time ('T...Z')
parse_any <- function(x) {
  # try m/d/Y
  d1 <- suppressWarnings(mdy(x))
  # try plain Y-m-d
  d2 <- suppressWarnings(ymd(x))
  # try ISO datetime
  d3 <- suppressWarnings(ymd_hms(x, quiet = TRUE))
  out <- coalesce(d1, d2, as_date(d3))
  as_date(out)
}

# ---- Parse & filter Jan–Aug by Completed Date ----
dat <- dat |>
  mutate(Date_for_Month = parse_any(Application.Completed.Date)) |>
  filter(!is.na(Date_for_Month)) |>
  filter(month(Date_for_Month) %in% 1:11)

# ---- Keep 2024/2025 and Accelerated (exclude Graduate/Master) ----
filtered_data <- dat |>
  filter(str_detect(Active.Term...Calculated, "2024|2025|2026")) |>
  filter(!str_detect(Active.Major...Calculated, regex("Graduate|Master", ignore_case = TRUE))) |>
  mutate(
    Year = if_else(str_detect(Active.Term...Calculated, "2024"), "2024", "2025", "2026"),
    Major_Short = if_else(
      as.character(Active.Major...Calculated) %in% names(all_abbreviations),
      all_abbreviations[as.character(Active.Major...Calculated)],
      as.character(Active.Major...Calculated)
    )
  ) |>
  group_by(Major_Short, Year) |>
  summarise(count = n(), .groups = "drop")

if (nrow(filtered_data) == 0) {
  stop("No rows after filtering. Confirm there are Jan–Aug completed dates and 2024/2025 in Active.Term...Calculated.")
}

# ---- Order majors by total across years (desc) ----
order_levels <- filtered_data |>
  group_by(Major_Short) |>
  summarise(total = sum(count), .groups = "drop") |>
  arrange(desc(total)) |>
  pull(Major_Short)

filtered_data <- filtered_data |>
  mutate(
    Major_Short = factor(Major_Short, levels = order_levels),
    Year = factor(Year, levels = c("2024","2025", "2026"))
  )

# ---- Plot ----
ggplot(filtered_data, aes(x = Major_Short, y = count, fill = Year)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  labs(
    title = "Accelerated Applications Completed (Jan–Aug) by Major\n(Source: Application Completed Date)",
    x = "Major",
    y = "Count",
    fill = "Year"
  ) +
  scale_fill_manual(values = okabe_ito, drop = FALSE) +
  theme_tufte(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.position = "top"
  )



```

### Application Start Distribution, year over year

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 8
#| out-width: 100%
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)

# Load CSV
data <- read.csv("C:\\Users\\Todd\\Dropbox\\Monthly Export\\Accelerated Applicants.csv")

# Convert Application Start Date to date format
data <- data %>%
  mutate(Application.Start.Date = as.Date(Application.Start.Date, format = "%m/%d/%Y"))

# Extract year from Start Term and filter for 2024 or 2025
data_filtered <- data %>%
  filter(grepl("2024|2025|2026", Active.Term...Calculated)) %>%
   mutate(
    Year = case_when(
      grepl("2024", Active.Term...Calculated) ~ "2024",
      grepl("2025", Active.Term...Calculated) ~ "2025",
      grepl("2026", Active.Term...Calculated) ~ "2026",
      TRUE ~ "Other"))

# Aggregate data: count application starts per day for each year
daily_counts <- data_filtered %>%
  group_by(Application.Start.Date, Year) %>%
  summarise(Count = n(), .groups = "drop")
library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove(Active.Major...Calculated, "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in "))

plot <- ggplot(data_filtered, aes(x = Application.Start.Date, fill = Year)) +
  geom_histogram(binwidth = 30, alpha = 0.6, position = "identity") +  # Semi-transparent overlap
  scale_color_brewer(palette = "Set2") +
  scale_x_date(date_breaks = "3 months", date_labels = "%m-%y") +  # Show full month and year
  scale_fill_manual(values = c("2024" = "#E69F00", "2025" = "#56B4E9", "2026" = "#009E73")) +
  labs(title = "Application Start Distribution (2024 - 2026)",
       x = "Application Start Date",
       y = "Number of Applications Started",
       fill = "Start Term Year") +
  theme_minimal(base_size=10) +
theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(plot)
```

### Cumulative Starts by year and program, small multiples

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center
#| out-width: "100%"
# Load required libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)

data <- read.csv("data/AcceleratedApps.csv")

# Clean the data
data <- data %>%
  mutate(Application.Start.Date = mdy(Application.Start.Date),  # Convert MM/DD/YYYY to Date format
         Year = year(Application.Start.Date),  # Extract Year
         Month = floor_date(Application.Start.Date, "month"),  # Keep Month
         Active.Major...Calculated = as.factor(Active.Major...Calculated)) %>%
  filter(!is.na(Application.Start.Date))  # Remove rows with NA values

# Check if Active.Term...Calculated exists before using it
if ("Active.Term...Calculated" %in% colnames(data)) {
  data <- data %>%
    mutate(Active.Term.Year = str_extract(as.character(Active.Term...Calculated), "\\d{4}"))  # Extract Year
} else {
  stop("Error: 'Active.Term...Calculated' column not found in the dataset.")
}

# 🔹 Filter for Graduate/Master's programs and select both 2024 & 2025 terms
data_filtered <- data %>%
  filter(grepl("Accelerated", Active.Major...Calculated, ignore.case = TRUE) &
           Active.Term.Year %in% c("2024", "2025", "2026"))  # Include both years

# Aggregate data by Month, Major, and Year, then compute cumulative sum
aggregated_data <- data_filtered %>%
  group_by(Month, Active.Major...Calculated, Active.Term.Year) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(Month) %>%
  group_by(Active.Major...Calculated, Active.Term.Year) %>%
  mutate(cumulative_count = cumsum(count))  # Compute cumulative sum

library(stringr)  # if not already loaded

# Simplify major names
aggregated_data <- aggregated_data %>%
  mutate(Major_Short = str_remove(Active.Major...Calculated, "Accelerated Bachelor's in ") %>%
                        str_remove("Graduate Certificate in ") %>% str_remove("Master"))

# Create the small multiples plot
plot <- ggplot(aggregated_data, aes(x = Month, y = cumulative_count, 
                            color = Active.Term.Year, 
                            group = Active.Term.Year)) +
  geom_point(size = 2) + 
  geom_line(size = 1) +
  scale_color_brewer(palette = "Set2") +  
  scale_x_date(date_breaks = "6 months", date_labels = "%b-%Y") +  
  labs(title = "Cumulative Application Starts by Major & Year",
       x = "Month",
       y = "Cumulative Applications Started",
       color = "Application Year") +  # Updated legend title
  facet_wrap(~ Major_Short, scales = "free_y") +  # Small multiples
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(plot)

```
Projection showing apps to date, compared against prior years, and projected total by end of year
```{r}
#| label: forecast_submit_only_oct2oct
#| warning: false
#| message: false
suppressPackageStartupMessages({
  library(dplyr); library(lubridate); library(stringr); library(tidyr)
})

# -------- Config --------
path <- "C:/Users/Todd/Dropbox/Daily Export/totalfunnelexport.csv"
cycle_start_month <- 10   # October
use_dedupe <- TRUE        # set FALSE to count all submits (no de-dup)
dedupe_key <- c("Element.ID")  # fallback order for person ID

# -------- Helpers --------
parse_any_date <- function(x) {
  d1 <- suppressWarnings(mdy(x))
  d2 <- suppressWarnings(ymd(x))
  d3 <- suppressWarnings(ymd_hms(x, quiet = TRUE))
  as_date(coalesce(d1, d2, d3))
}
# shift so Oct 1 behaves like Jan 1 (for day-in-cycle calc)
shift_to_cycle <- function(d, start_month = 10) d %m-% months(start_month - 1)
# label cycles by their start year (e.g., 2025 covers 2025-10-01..2026-09-30)
cycle_label_from_date <- function(d, start_month = 10) if_else(month(d) >= start_month, year(d), year(d) - 1)

# -------- Load --------
df <- read.csv(path, stringsAsFactors = FALSE, check.names = TRUE)

# Ensure the Submit column exists exactly as an R-safe name
if (!"Application.Submit.Date" %in% names(df)) {
  stop("Column 'Application.Submit.Date' not found. Run names(df) to confirm; 
       with check.names=TRUE it must have dots, not spaces.")
}

# Build working frame using SUBMIT ONLY
dat <- df %>%
  mutate(
    SubmitDate = parse_any_date(Application.Submit.Date)
  ) %>%
  filter(!is.na(SubmitDate)) %>%              # only real submits
  mutate(
    cycle_label   = cycle_label_from_date(SubmitDate, cycle_start_month),
    day_in_cycle  = yday(shift_to_cycle(SubmitDate, cycle_start_month)),
    person_key    = coalesce(!!!rlang::syms(dedupe_key))  # School.ID then Email
  ) %>%
  filter(!is.na(day_in_cycle))

# Optional de-dup: keep ONE submit per person per cycle (earliest submit)
if (use_dedupe) {
  dat <- dat %>%
    group_by(cycle_label, person_key) %>%
    slice_min(order_by = SubmitDate, n = 1, with_ties = FALSE) %>%
    ungroup()
}

# Identify cycles
cycles_available <- sort(unique(dat$cycle_label))
current_cycle <- cycle_label_from_date(Sys.Date(), cycle_start_month)
hist_cycles <- setdiff(cycles_available, current_cycle) |> tail(2)

if (length(hist_cycles) == 0) {
  stop("No prior cycles available to build a seasonal curve. Check that Application.Submit.Date is populated in earlier cycles.")
}

hist_data    <- dat %>% filter(cycle_label %in% hist_cycles)
current_data <- dat %>% filter(cycle_label == current_cycle)

# -------- Historical share curve (Oct→Oct) --------
hist_daily <- hist_data %>%
  count(cycle_label, day_in_cycle, name = "submits") %>%
  group_by(cycle_label) %>%
  complete(day_in_cycle = 1:366, fill = list(submits = 0)) %>%
  arrange(day_in_cycle) %>%
  mutate(total = sum(submits),
         cum   = cumsum(submits),
         share = if_else(total > 0, cum/total, 0)) %>%
  ungroup()

share_curve <- hist_daily %>%
  group_by(day_in_cycle) %>%
  summarise(
    mean_share = mean(share, na.rm = TRUE),
    p25_share  = quantile(share, 0.25, na.rm = TRUE),
    p75_share  = quantile(share, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

# Today's position in the cycle
today_day_in_cycle <- yday(shift_to_cycle(Sys.Date(), cycle_start_month))
curve_today <- share_curve %>% filter(day_in_cycle <= today_day_in_cycle) %>% slice_tail(n = 1)

# Guard rails if very early in cycle
if (nrow(curve_today) == 0 || is.na(curve_today$mean_share) || curve_today$mean_share <= 0) {
  curve_today <- tibble::tibble(mean_share = 0.05, p25_share = 0.03, p75_share = 0.08)
}

# -------- Projection (Submit-based) --------
submits_to_date <- nrow(current_data)
mean_share <- max(0.01, curve_today$mean_share)
p25_share  <- max(0.005, curve_today$p25_share)
p75_share  <- max(0.005, curve_today$p75_share)

projected_total_mean <- ceiling(submits_to_date / mean_share)
projected_total_low  <- ceiling(submits_to_date / p75_share)  # optimistic (faster season)
projected_total_high <- ceiling(submits_to_date / p25_share)  # conservative (slower season)

# -------- Diagnostics + Result --------
diag_tbl <- tibble::tibble(
  date_column_used = "Application.Submit.Date",
  deduped = use_dedupe,
  cycles_available = paste(cycles_available, collapse = ", "),
  history_cycles_used = paste(hist_cycles, collapse = ", "),
  current_cycle = current_cycle,
  submits_to_date_current_cycle = submits_to_date
)
print(diag_tbl)

tibble::tibble(
  Cycle                 = current_cycle,
  Through               = as.character(Sys.Date()),
  Submits_to_Date       = submits_to_date,
  Share_Today_Mean      = round(mean_share, 3),
  Projected_Final_Mean  = projected_total_mean,
  Projected_Range       = sprintf("%s–%s", projected_total_low, projected_total_high)
)

